All samples yielded sufficient biofilm growth and starch incorporation to be 
included in the analysis (Figure \@ref(fig:microscope-fig)), resulting in a total of 48 biofilm samples (two plates of 24),
45 of which were used for analysis (three samples were set aside for later
 analysis).
Most control samples contained no starch granules, while some contained negligible
quantities (see Supplementary Material).

## No amylase activity detected in the model

```{r amylase-activity}
# table of results reported in units amylase per mL enzyme (but let's be honest,
  # ...it doesn't really matter what the unit is. No activity is no activity)
cols <- c("1", "2", "3")  
rows <- c("S1", "S2", "S3", "B1", "B2", "B3", "B4", "B5", "BT1", "BT2", "BT3")
plt1_ph1_result <- rbind(sal1_ph1, bmm1_ph1)
  plt1_ph1_result[plt1_ph1_result < 0] <- 0
  
amylase_results <- rbind(sal1_ph1, sal1_ph2, sal2_ph1, sal2_ph2, bmm1_ph1,
                         bmm1_ph2, bmm2_ph1, bmm2_ph2)
amylase_results[amylase_results < 0] <- NA
positive_amylase <- na.omit(amylase_results)
```

No $\alpha$-amylase activity was detected in any of the artificial
saliva samples from any of the days that were sampled. Only positive controls
(saliva)
contained amylase activity that could be detected in the assay, ranging from
`r paste(signif(range(positive_amylase),3), collapse = " to ")` U/mL enzyme 
(full results can be found in the Supplementary Material).
The results are not comparable to other studies presenting $\alpha$-amylase activity
levels in humans, as the unit definition may differ; 
however, they are sufficient to show that there is no activity in the model.

## Treatment type had minimal effect on biofilm growth

A one-way ANOVA suggests that the type of starch used during the biofilm 
growth period had a minimal effect on the growth of the biofilm 
(expressed as total dry weight of the sample), F(`r paste(treat_df[1], treat_df[2], sep = ", ")`)
= `r treat_f.stat`, p = `r treat_p`.
A summary of sample weights is available in Table \@ref(tab:anova-tbl).

## Starch counts

<!-- starch counts in the solutions -->

```{r solution-count-out}
sol_tbl <- sol_comb_out %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")"))) 

sol_tbl[1,3] <- NA
```

<!-- starch counts in the samples -->

```{r sample-count}
samp_counts_per <- samp_counts %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(signif(x / total, 3)), 
                         ")")))

# combine samp_counts_per with samp_sd_wide
samp_counts_tbl <- samp_counts_per %>%
  bind_cols(select(samp_sd_wide, c(s_sd, m_sd, l_sd, total_sd)), .name_repair = "unique") %>%
  select(treatment, starch, s, s_sd, m, m_sd, l, l_sd, total, total_sd)

samp_counts_tbl[1,3] <- NA
```

It was not possible to differentiate between potato and 
wheat starches smaller than ca. 10 $\mu$m. These were counted as wheat,
as we assumed that the majority of the small granules were wheat. We make this 
assumption based on the counts of small starches in the wheat-only and potato-only
solutions. Of the combined amount of small starches in these two solutions, 
`r paste0(signif((sol_comb_out$s[5] / sum(sol_comb_out$s[4], sol_comb_out$s[5])) * 100, 3), "%")`
are from wheat.
<!-- count of small granules from wheat solution divided by the sum of small granules 
from the potato and wheat solutions -->

<!-- absolute solution counts compared to absolute sample counts. does exposure
to higher quantities of starches mean more incorporated starches? -->
The separate wheat and potato solutions were made with a 0.25% (w/v) starch
concentration, while the mixed-starch solution was made with 0.25% (w/v) of each
starch, with a total concentration of 0.50% (w/v).
The mixed treatment had the highest absolute count of starch granules in solution (mean = `r signif(sol_comb_row$total, 3)`),
while the biofilms exposed to the wheat solution preserved the greatest number of
granules 
(mean = `r signif(subset(samp_counts, treatment == "wheat")$total, 3)`). 
The potato treatment had the lowest absolute counts in both the solution 
(`r signif(subset(sol_comb_out, treatment == "potato")$total, 3)`)
and in the biofilm samples 
(`r signif(subset(samp_counts, treatment == "potato")$total, 3)`) 
(Tables \@ref(tab:solution-count-tbl) and \@ref(tab:sample-count-tbl)).

### Proportion of available starches incorporated in samples

```{r sample-perc}
prop_wide <- prop %>%
  ungroup() %>%
  pivot_wider(id_cols = c(treatment, starch), 
              names_from = size, 
              values_from = prop) %>%
  add_row(prop_comb_row, .after = 2)
```

The proportion of total starches from the solutions that were incorporated 
into the samples ranged from 
`r paste(min(perc_tbl$total), max(perc_tbl$total), sep = " to ")`,
with potato granules being more readily incorporated than wheat in both the
separated- and mixed-treatment samples
(Table \@ref(tab:sample-prop-tbl)). 
There is an inverse relationship between the absolute starch count in the solutions 
and the proportional incorporation of starches in the biofilm samples, i.e., potato 
had the lowest absolute count in solutions, but the highest proportional 
incorporation, and vice versa for the mixed treatment.

<!-- mixed treatments -->
Wheat incorporation was most affected in the mixed-treatment samples, with only
`r paste0(perc_tbl$total[2])` of the total available starches being incorporated into
the sample, compared to
`r paste0(perc_tbl$total[4])`
in the separated wheat treatment.

### Size ratios differ between solutions and samples

<!-- size proportions -->
<!-- pie charts -->

<!-- overall size trends -->

```{r size-trends}
sol_size_prop <- sol_comb_out %>%
  group_by(treatment, starch) %>%
  #select(!total) %>%
  mutate(across(where(is.numeric), function(x) x / total)) %>%
  mutate(type = "solution")

samp_size_prop <- samp_counts %>%
  group_by(treatment, starch) %>%
  #select(!total) %>%
  mutate(across(where(is.numeric), function(x) x / total)) %>%
  mutate(type = "sample")

size_diff <- sol_size_prop %>%
  full_join(samp_size_prop) %>%
  pivot_longer(cols = c(s,m,l,total), names_to = "size", values_to = "count") %>%
  pivot_wider(names_from = type, values_from = count) %>%
  mutate(diff = (sample - solution) * 100) %>%
  select(!c(solution, sample)) %>%
  pivot_wider(names_from = size, values_from = diff) %>%
  select(!total)
```


Overall, medium starch granules had a higher mean rate of incorporation 
(`r scales::percent(mean(prop_wide$m, na.rm = T), 0.001)`) 
than small 
(`r scales::percent(mean(prop_wide$s, na.rm = T), 0.001)`) 
and large 
(`r scales::percent(mean(prop_wide$l, na.rm = T), 0.001)`) 
starch granules across all treatments, while large potato starches had the lowest 
rate of incorporation across all treatments.

<!-- separated and mixed treatments -->
The difference in incorporation between the size categories resulted in a change
in size ratios between the original starch solutions and the extracted samples.
Large potato granules (> 20 $\mu$m) were most affected, with a 
`r paste0(abs(signif(size_diff$l[4], 3)), "%")` 
decrease in relative abundance in the potato-only treatment, and a
`r paste0(abs(signif(size_diff$l[1], 3)), "%")` 
decrease in mixed treatments. Medium granules increased in relative abundance 
across all samples, while small granules decreased in wheat treatments and 
increased in potato treatments 
(Figure \@ref(fig:ratio-plots)). 

<!-- Proportion and size ratios are the most important points to be made -->

### Biofilm weight correlated positively with extracted starch counts

```{r cor, echo=FALSE}
# prepare cor output for reporting
cor_r <- signif(count_cor$estimate, 3)
cor_ci <- signif(count_cor$conf.int, 3)
cor_p <- if(count_cor$p.value > 0.001){
  paste("=", signif(count_cor$p.value, 3))
} else {
  paste("< 0.001")
}

cor2_r <- signif(starch_cor$estimate, 3)
cor2_ci <- signif(starch_cor$conf.int, 3)
cor2_p <- if(starch_cor$p.value > 0.001){
  paste("=", signif(starch_cor$p.value, 3))
} else {
  paste("< 0.001")
}
# Convert r correlation result to text
direct_cor <- ifelse(cor_r < 0, "negative", "positive")
strength_cor <- if(cor_r >= 0.8){
  "very strong"
} else if(cor_r < 0.8 & cor_r >= 0.6){
  "strong"
} else if(cor_r < 0.6 & cor_r >= 0.4){
  "moderate"
} else if(cor_r < 0.4 & cor_r >= 0.2){
  "weak"
} else{
  "very weak"
}
```

Pearson's *r* suggests a 
`r strength_cor` `r direct_cor` 
correlation between the total weight of the biofilms and the total starch count 
(standardised by z-score) extracted from the samples across treatments, 
*r* = `r cor_r`, 
90%CI[`r paste(cor_ci[1], cor_ci[2], sep = ", ")`],
p `r cor_p`
(Figure \@ref(fig:cor-plot)A). 

The same test was applied to total biofilm weight and starch count per mg 
calculus (also standardised by z-score), resulting in a weak positive correlation, 
*r* = `r cor2_r`, 
90%CI[`r paste(cor2_ci[1], cor2_ci[2], sep = ", ")`],
p `r cor2_p`
(Figure \@ref(fig:cor-plot)B).
