---
title: Supplementary Materials
output: 
  html_document:
    keep_md: true
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
raw_data <- read_csv(here("analysis/data/raw_data/starch_counts.csv"))
```

## R Session info

```{r}
print(sessionInfo(), locale = FALSE)
```


## Metadata for raw data files

Counts represent the absolute number of starches counted on a slide

**starch_counts.csv**

| variable | description |
|----|----|
| sample | Sample number. |
| plate | Plate number that the sample came from. |
| row | Which row on the plate the sample came from. |
| s | Small starch count. |
| m | Medium starch count. |
| l | Large starch count. |
| total | Sum of s, m, and l. |
| treatment | Treatment solution to which the samples were exposed. |
| starch | Type of starch that was counted. |
| weight | Weight of the biofilm sample. |
| vol | Total volume of EDTA in which the sample was dissolved. |
| portion_slide | Proportion of the microscope slide that was counted. Total transects on slide divided by counted transects. |

**solution_counts**

| variable | description |
|----|----|
| solution | Type of starch in solution. |
| concentration | Concentration (%w/v) of starch in solution. | 
| vol_slide | Volume of solution added to slide. |
| vol_total | Total volume of solution in aliquot. |
| portion_slide | Proportion of slide that was counted. Total transects on slide divided by counted transects. |
| slide | Slide number. |
| starch | Starch type counted. |
| s | Small starch count. |
| m | Medium starch count. |
| l | Large starch count. |
| total | Sum of s, m, and l. |

## Amylase activity

Amylase activity in U/mL enzyme, where a U is mg maltose released from starch in six minutes at 36 &deg;C.

Tables containing the amylase activity results for both plates and photometric
readings.

```{r amylase-activity-results, message=FALSE}
# table of results reported in units amylase per mL enzyme (but let's be honest,
  # ...it doesn't really matter what the unit is. No activity is no activity)
source(here("analysis/scripts/amylase-assay.R"))
cols <- c("1", "2", "3")  
rows <- c("S1", "S2", "S3", "B1", "B2", "B3", "B4", "B5", "BT1", "BT2", "BT3")
plt1_ph1_result <- rbind(sal1_ph1, bmm1_ph1)
rownames(plt1_ph1_result) <- rows
plt1_ph2_result <- rbind(sal1_ph2, bmm1_ph2)
rownames(plt1_ph2_result) <- rows
plt2_ph1_result <- rbind(sal2_ph1, bmm2_ph1)
rownames(plt2_ph1_result) <- rows
plt2_ph2_result <- rbind(sal2_ph2, bmm2_ph2)
rownames(plt2_ph2_result) <- rows
```

```{r amylase-activity-tables}
plt1_ph1_result
plt1_ph2_result
plt2_ph1_result
plt2_ph2_result
```

## Control samples

```{r}
raw_data %>%
  filter(treatment == "control") %>%
  select(!c(vol, portion_slide, s, m, l))
```

Only the total starch count was considered for control samples, as size was
deemed irrelevant.

## Count corrections

Slide transects were calculated by counting the number of transects on the 
cover slip under the microscope. This was done by starting in the bottom-left
corner, and counting the total number of full fields-of-view across the cover 
slip to the bottom-right corner. The total number of transects was 29 (verified
multiple times).

Solution counts were extrapolated to the quantity in a 1 ml solution with
the following formula:

$$ \text{corrected count} = \text{raw count} \times \frac{\text{total} \space \text{transects}}{\text{counted} \space \text{transects}} \times 100 \mu l $$

Sample counts were extrapolated using the following formula:

$$ \text{Corrected count} = \text{raw count} \times (\text{portion of slide})^{-1} \times \frac{V_{sample}}{V_{slide}} $$

## Microscope images

```{r img-1, echo=FALSE, fig.cap="Image of starch granules extracted from a calculus sample"}

```

