---
title: Supplementary Materials
output: 
  html_document:
    keep_md: true
    df_print: paged
---

```{r setup}
library(here)
library(tidyverse)
raw_data <- read_csv(here("analysis/data/raw_data/starch_counts.csv"))
```


## Metadata

**raw_counts**
<!-- insert data type for each column -->
Counts represent the absolute number of starches counted on a slide

sample: Sample number.
plate: Plate number that the sample came from.
row: Which row on the plate the sample came from.
s: Small starch count.
m: Medium starch count.
l: Large starch count.
total: Sum of s, m, and l.
treatment: Treatment solution to which the samples were exposed.
starch: Type of starch that was counted
weight: Weight of the biofilm sample
vol: Total volume of EDTA in which the sample was dissolved.
portion_slide: Proportion of the microscope slide that was counted.

## Amylase activity

Tables containing the amylase activity results for both plates and photometric
readings.

```{r amylase-activity-results}
# table of results reported in units amylase per mL enzyme (but let's be honest,
  # ...it doesn't really matter what the unit is. No activity is no activity)
cols <- c("1", "2", "3")  
rows <- c("S1", "S2", "S3", "B1", "B2", "B3", "B4", "B5", "BT1", "BT2", "BT3")
(plt1_ph1_result <- rbind(sal1_ph1, bmm1_ph1))
#  plt1_ph1_result[plt1_ph1_result < 0] <- 0
(plt1_ph2_result <- rbind(sal1_ph2, bmm1_ph2))
#  plt1_ph2_result[plt1_ph2_result < 0] <- 0
(plt2_ph1_result <- rbind(sal2_ph1, bmm2_ph1))
  #plt2_ph1_result[plt2_ph1_result < 0] <- 0
(plt2_ph2_result <- rbind(sal2_ph2, bmm2_ph2))
  #plt2_ph2_result[plt2_ph2_result < 0] <- 0
```

## Control samples

```{r}
raw_data %>%
  filter(treatment == "control") %>%
  select(!c(vol, portion_slide, s, m, l))
```

Only the total starch count was considered for control samples, as size was
deemed irrelevant.

## Count corrections

Sample counts were extrapolated using the following formula:

$$ \text{Corrected count} = \text{raw count} \times (\text{portion of slide})^{-1} \times \frac{V_{sample}}{V_{slide}} $$

Solution counts were extrapolated to the quantity in a 1 ml solution with
the following formula:

$$ \text{corrected count} = \text{raw count} \times \frac{29}{3} \times 100 \mu l $$

## Functions
